---
title: "p8106_hw6_gw2442"
output: github_document
date: "2022-11-28"
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(broom)
library(patchwork)
library(dplyr)
set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 2

#### Importing and describing raw data: 

```{r}
homicide_data = read_csv(file = "./data/homicide-data.csv") %>%
  janitor::clean_names()
skimr::skim(homicide_data)
```

The CSV file was downloaded from the GitHub repository, imported, and cleaned. The dataset `homicide_data` has 52,179 rows and 12 columns. The dataset contains information on the homicide ID number, reported date, the victim's first and last name, race, age, sex, location of homicide (including city, state, longtidue, and latitude), and disposition of the case. While there are 0 missing variables for variables `uid`, `victim_last`, `victim_first`, `victim_race`, `victim_age`, `victim_sex`, `city`, `state`, `disposition`, and `reported date`, there are 60 missing variables for `lat` and `lon`. 


#### Creating variable `city_state`:

```{r}
homicide_data =
  homicide_data %>%
  mutate(
    city_state = as.character(paste(city, state, sep = "_")))
```


#### Creating variable `unsolved_homicide` and tidying data: 

The variable `solved_homicide` was created. Those with the disposition "Closed without arrest" or "Open/No arrest" were categorised as an unsolved homicide and were therefore given a value of 0 in the variable `solved_homicide`. Those with the disposition "Closed by arrest" were categorised as a solved homicide and were therefore given a value of 1 in the variable `solved_homicide`. 

The cities Dallas, Phoenix, Kansas City, and Tulsa were omitted from the dataset, as well as any race entries other than White or Black and any victim age's that were non-numeric. 

```{r}
homicide_data = 
  homicide_data %>%
  mutate(solved_homicide = ifelse(disposition == "Closed by arrest", 1, 0)) %>%
  filter(city != "Dallas", 
         city != "Phoenix", 
         city != "Kansas City", 
         city != "Tulsa") %>%
  filter(victim_race == "White" | victim_race == "Black") %>%
  mutate(victim_age = as.numeric(victim_age)) %>%
  drop_na(victim_age)
```

#### Logistic regression on Baltimore, MD